---
title: "648 Lab 3"
author: "gucci fan"
date: "2024-11-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Given a dataset of Flickr photographs with a predicted classification of either nature=TRUE or nature=FALSE, and latitude and longitude locations of the photographs that correspond with a NLCD land cover/use type, I am hypothesizing that the spatial distribution of the nature=FALSE data points align with the NLCD land use categories of 21: Developed Open Space, 22: Developed Low Intensity, 23: Developed Medium Intensity, and 24: Developed High Intensity. I also hypothesize that the spatial pattern of nature=TRUE photographs will be more similar to the distribution of all other NLCD categories, including Open Water, Mixed Forest, and Woody Wetlands, ie. "nature". 

```{r}
library(readr)
library(ggplot2)
library(ggthemes)
library(gganimate)
library(foreign)
library(dplyr)
library(rnaturalearth)
library(rnaturalearthdata)
library(gifski)
library(lubridate)
#library(tidyverse)
library(wesanderson)
library(sf)

```

```{r}
MichFlickr <- read.csv("MichiganFlickr.csv")
```

```{r}
MichFlickr$date <- as.POSIXct(MichFlickr$dateupload, origin="1970-01-01")
MichFlickr$date <- as.Date(format(MichFlickr$date, format="%Y-%m-%d"))
MichFlickr$year <- year(MichFlickr$date)
MichFlickr$month <- month(MichFlickr$date, label = TRUE)
MichFlickr$day <- day(MichFlickr$date)
MichFlickr$count<- 1
MichFlickr$Nature<- MichFlickr$predict_Na > 0.6
```

To investigate the spatial distribution of photos in relation to their classified land use, as well as compare that to whether the photo was classified as nature or not nature, I am first sorting by land use type, and dropping data points where land use was 0 (NA).

```{r}
unique(MichFlickr$Landuse)
sum(MichFlickr$Landuse == 41)

landuse_freq <- MichFlickr %>% filter(Landuse != 0) %>%
  group_by(Landuse) %>%
  summarise(total_photos=sum(count)) %>%
  na.omit()
head(landuse_freq)

landuse_labels <- c("Open Water", "Developed, Open Space", "Developed, Low Intensity", "Developed, Medium Intensity", "Developed, High Intensity", "Barren Land (Rock/Sand)", "Deciduous Forest", "Evergreen Forest", "Mixed Forest", "Shrub/Scrub", "Grassland/Herbaceous", "Pasture/Hay", "Cultivated Crops", "Woody Wetlands", "Emergent Herbaceous Wetlands")

landuse_freq <- landuse_freq %>%
  mutate(Proportion = total_photos / sum(total_photos))

ggplot(landuse_freq, aes(x = "", y = Proportion, fill = factor(Landuse))) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar("y", start = 0) + 
  labs(fill = "Land Use Type", y = NULL, x = NULL, 
       title = "Proportions of Land Use Types") +
  theme_void() +  
  geom_text(aes(label = scales::percent(Proportion, accuracy = 0.01)), 
            position = position_stack(vjust = 0.5), size=3) +
  scale_fill_manual(values = c("#486da1", "#e1cdce", "#dc9881", "#f10100", "#ab0101", "#b3afa4", "#6ca966", "#1d6533", "#bdcc93", "#d1bb82", "#edeccc", "#ded83e", "#ae7229", "#bbd7ed", "#71a4c1"),
                    labels = landuse_labels)

```

From this pie chart, the pink to red slices represent developed land use types and they are about half of the entire dataset. In the non-developed slices, open water, deciduous forest, and cultivated crops make up the most significant "nature" type land uses.

I will compare the proportions of photographs with "nature" vs "non-nature" predicted categories, as well as "natural" vs "developed" categories of NLCD land uses for those same photos.

```{r}

count_developed <- landuse_freq %>%
  filter(Landuse %in% c(21, 22, 23, 24)) %>% 
  summarise(total_photos = sum(total_photos)) 

developed_proportion <- count_developed$total_photos / sum(landuse_freq$total_photos)

developed_proportion
```


A little over half of the photos, or 55.73%, were taken at a location classified as one of the "developed" NLCD land use types - 21: Developed open space, 22: Developed low intensity, 23: Developed medium intensity, or 24: Developed high intensity. That means 44.27% of photographs were taken at "non-developed" land uses sites. I wanted to compare this with the proportion of nature vs non-nature photographs (classified independently of land use). Basically, do people take photos of nature at locations classified as non-developed? What would cause discrepancies?

```{r}
nature_false <- sum(MichFlickr$Nature == FALSE)
nature_true <- sum(MichFlickr$Nature == TRUE)
proportion_nature <- nature_true / sum(nature_true + nature_false)
proportion_nature
```
The proportion of nature to non-nature is much lower than expected, with only 21% of photos being classified as nature.

```{r}
states <- map_data("state")
mich <- subset(states, region == "michigan")
mich <- states %>%
  filter(region == "michigan")
counties <- map_data("county")
mich_county <- subset(counties, region == "michigan")

base_mich <- ggplot(data=mich) +
  geom_polygon(mapping = aes(x = long, y = lat, group = group), fill = "gray90", color = "black") +
  coord_fixed(1.3)

base_mich
```

I am building a base Michigan map from map_data to plot spatial patterns of the photograph data over it.

```{r}
landuse_map <- c(
  "11" = "Open Water", 
  "21" = "Developed, Open Space", 
  "22" = "Developed, Low Intensity", 
  "23" = "Developed, Medium Intensity", 
  "24" = "Developed, High Intensity", 
  "31" = "Barren Land (Rock/Sand)", 
  "41" = "Deciduous Forest", 
  "42" = "Evergreen Forest", 
  "43" = "Mixed Forest", 
  "52" = "Shrub/Scrub", 
  "71" = "Grassland/Herbaceous", 
  "81" = "Pasture/Hay", 
  "82" = "Cultivated Crops", 
  "90" = "Woody Wetlands", 
  "95" = "Emergent Herbaceous Wetlands"
)

MichFlickr$LanduseNames <- factor(MichFlickr$Landuse, levels = names(landuse_map), labels = landuse_map)
```

I added the land use names to the dataset to be interpreted better.

```{r}
michstatic <- base_mich +
  geom_point(data = MichFlickr, aes(x = longitude, y = latitude, color = LanduseNames), size = 0.5) + 
  scale_color_manual(values = c("#486da1", "#e1cdce", "#dc9881", "#f10100", "#ab0101", "#b3afa4", "#6ca966", "#1d6533", "#bdcc93", "#d1bb82", "#edeccc", "#ded83e", "#ae7229", "#bbd7ed", "#71a4c1")) +
  labs(title = "Flickr Photos by Land Use Type in Michigan",
       color = "Land Use Type") +
  guides(color = guide_legend(override.aes = list(size = 3)))+
  theme_minimal()

michstatic
```
This is a static version of the distribution of land use types illustrated via Flickr photos. I used the colors of the NLCD land use classifications.

```{r}
animateMich <- MichFlickr %>% 
  filter(Landuse != 0)

# p2 <- base_mich +
#     geom_point(data = animateMich, aes(x = longitude, y = latitude, color = LanduseNames), size = 1) +
#     transition_states(LanduseNames, transition_length = 3, state_length = 3) +
#     labs(title = "Flickr Photos by Land Use Type in Michigan", subtitle = "Land Use: {closest_state}", color = "Land Use Type")
# 
# p2

p3 <- michstatic + 
    transition_states(LanduseNames, transition_length = 3, state_length = 3) +
    labs(title = "Flickr Photos by Land Use Type in Michigan", subtitle = "Land Use: {closest_state}", color = "Land Use Type")

p3
   
```

```{r}
animate(p3, duration = 10, fps = 20, width = 800, height = 600, renderer = gifski_renderer("landuse_nlcd_animation.gif"))
```

I can visualize the distribution of land use with an animation iterating through the types of land use associated with the spatial location of the photos. Comparing to a known land use map from NLCD, it is interesting to see how land use from Flickr photographs are distributed - some are obvious like 11: Open Water being mostly concentrated around the coastlines, and 24: Developed, High Intensity is concentrated around the metro Detroit area. 

Now I am categorizing the data by nature vs. non-nature, and comparing the spatial patterns there to splitting the data by developed vs. non-developed NLCD land use.

```{r}
michnaturestatic <- base_mich +
  geom_point(data = MichFlickr, aes(x = longitude, y = latitude, color = Nature), size = 0.5) + 
  scale_color_manual(values = c("TRUE" = "#1d6533", "FALSE" = "#ab0101")) +
  labs(title = "Flickr Photos by Nature Classification in Michigan") +
  guides(color = guide_legend(override.aes = list(size = 3)))+
  theme_minimal()

michnaturestatic
```
```{r}
MichFlickr$Developed <- MichFlickr$Landuse %in% c(21, 22, 23, 24)

michdevelopedstatic <- base_mich +
  geom_point(data = MichFlickr, aes(x = longitude, y = latitude, color = Developed), size = 0.5) + 
  scale_color_manual(values = c("TRUE" = "#ab0101", "FALSE" = "#1d6533")) +
  labs(title = "Flickr Photos by Developed vs Non-Developed Land Use") +
  guides(color = guide_legend(override.aes = list(size = 3)))+
  theme_minimal()

michdevelopedstatic
```

It could perhaps be of interest to animate this data by making a gif that flips back and forth between the developed/non-developed and nature/non-nature maps.

Some takeaways from this data are:
- Nature=TRUE occurs more in the northern half of the state, although more photographs in general are concentrated in the Lower Peninsula.
- Perhaps in the metro Detroit area, both datasets trend towards a higher concentration of non-nature and developed.
- Developed=TRUE shows up as distinct clusters where cities and towns are, while Nature=FALSE does not follow a distinct pattern.

Some conclusions could be that a lot more photos are showing up as Nature=FALSE (roughly 80% of the photographs), than Developed=TRUE (roughly 56%) because despite being at a land use site such as Open Water or Deciduous Forest, the photograph could still be categorized as Nature=FALSE perhaps due to the subject of the photo being of people and not environment-focused. This is interesting to note if someone wanted to know the content of photographs - do Flickr users gravitate more towards posting photographs of people or non-natural objects, even when in a natural environment setting?

Lab question 2:
To counteract "noisy" data, ie. plotting the individual occurrences of photos, binning can be done to group spatially and/or temporally adjacent points together. Similar to giving a histogram wider bars, the data points can be mapped by wider time intervals or bigger spatial cells to smooth out this data. Furthermore, setting a threshhold to exlclude single outliers, such as 1 photograph in a spatial cell of 5km x 5km, can also reduce noise as these outliers are likely not crucial to the overall pattern.

